---
title: "Assignment 2B"
author: "Brian Roepke"
date: "Sept 13, 2020"
output:
  html_document:
    df_print: paged
---
## Project Setup

```{r}
# clear all objects includes hidden objects from Global Environment
rm(list = ls(all.names = TRUE)) 
ls()
```

**Write a function compute_s_n that for any given n,**

 * Computes the sum: Sn = 1^2 + 2^2 + 3^2 + .. n^2
 * Report the value of the sum when n = 10. Eg: compute_s_n(10). You can enclose this in the print function
 * Hint: (1:5)^2 returns: 1 4 9 16 25. This squares each number from 1 to 5. The sum is 55 (1 + 4 +9 +16 +25)

```{r}
n <- 10

compute_s_n <- function(n){
  total = sum((1:n)^2)
}
```


```{r}
# validate the function returns 55 for the number 5
print(compute_s_n(5))

# run the funcation with the number 10
print(compute_s_n(10))
```

**Define an empty numerical vector s_n of size 25 using s_n <- vector(“numeric”, 25)**

 * store in the results of S1, S2,…S25 using a for-loop
 
```{r}
m <- 25
s_n <- vector("numeric", length=m)  # create an empty vector

print ('s_n before assignment')
print(s_n)

## Add your for loop
for(i in 1:m){
 s_n[i] <- i 
}

# uncomment once you've added the code for the for-loop
print('s_n contents using for-loop')
print(s_n)
```
 
**Repeat the above using sapply**

You will need to create a function, call it m_funct
 
```{r}
m <- 25
s_n <- vector("numeric", length=m) 

print ('s_n before assignment')
print(s_n)

# add code for m_funct
 m_funct <- function(x){
   x
 }

# uncomment once you've defined m_funct
s_n = sapply(1:25, m_funct)
print ('s_n after assignment')
print(s_n)
```




 