---
title: "Assignment 2A"
author: "Brian Roepke"
date: "Sept 13, 2020"
output:
  html_document:
    df_print: paged
---
## Project Setup

```{r}
# clear all objects includes hidden objects from Global Environment
rm(list = ls(all.names = TRUE)) 
ls()
```


## Prerequsite

```{r}
library(dslabs) # load the package library

# Load the US murders dataset from the library
data(murders)
```

## Part 1

 * Use function str to examine the structure of murders.
 * Use head and tail functions to display the top/bottom records.
 * Use the accessor $ to extract state abbreviations and assign them to an object a. Show the class of a?
 * Use square brackets to extract state abbreviations and assign them to the object b. Use the identical function to determine if a and b are the same (hint: get usage help: ?identical).
 * Use levels to show the regions (a factor) then use nlevels function to show the number of regions.
 * Use the table function to create a table of states per region
 
```{r}
str(murders)
```
 
```{r}
head(murders)
```
```{r}
tail(murders)
```
 
```{r}
a <- murders$abb

class(a)
a
```

```{r}
b <- murders[['abb']]

class(b)
```

```{r}
identical(a,b)
```

```{r}
levels(murders$region)
nlevels(murders$region)
```

```{r}
table(murders$region)
```

## Part 2

 * Use the c (concatenate) function to create a vector called ‘temp’ containing high temperatures for Beijing, Lagos, Paris, Rio de Janeiro, San Juan, and Toronto with values: 35, 88, 42, 84, 81, and 30. Then print temp and show its class type.
 * Create a vector called ‘city’ with the city names from temp (hint: use names with the c function). Then print city and show its class type.
 * Use the [ and : operators to access the temperature of the first 3 cities of ‘temp’.
 * Use the [ operator to access the temperature of Paris and San Juan.
 * Use named index to get the temperature of Beijing
 * Create a sequence called s with numberic values 12,13,14, …,73. Print it and show its class type.
 * Create the sequence: a = seq(1, 10, 0.5). Then print it and show its class type.
 * Create the sequence: b = seq(1, 10). Then print it and show its class type.
 * Define the following vector: x = c(“1”, “3”, “5”) and coerce it to get integers. Show class of x
 
```{r}
temp <- c('Beijing' = 35, 'Lagos' = 88, 'Paris' = 42, 'Rio de Janeiro' = 84, 'San Juan' = 81,'Toronto' = 30)
# temp <- c(35, 88, 42, 84, 81, 30)
temp
class(temp)
```

```{r}
city <- names(temp)
city
class(city)
```

```{r}
temp[1:3]
```

```{r}
temp[3]
temp[5]
```

```{r}
temp['Beijing']
```

```{r}
s <- seq(12,73)
s
class(s)
```

```{r}
a = seq(1, 10, 0.5)
a
class(a)
```

```{r}
b = seq(1, 10)
b
class(b)
```

```{r}
x = c("1", "3", "5")
x
class(x)
x <- as.numeric(x)
class(x)
```

## Part 3

 * Create a data frame from the following 4 given vectors and display the data frame:
name: Anastasia, Dima, Katherine, James, Emily, Michael, Matthew, Laura, Kevin, Jonas
score: 12.5, 9, 16.5, 12, 9, 20, 14.5, 13.5, 8, 19
attempts: 1, 3, 2, 3, 2, 3, 1, 1, 2, 1
qualify: yes, no, yes, no, no, yes, yes, no, no, yes
 * Get the structure of the above data frame.
 * Get the statistical summary and nature of the data of the above data frame.
 * Extract 3rd and 5th rows with 1st and 3rd columns from the above data frame.
 
```{r}

name <- c('Anastasia', 'Dima', 'Katherine', 'James', 'Emily', 'Michael', 'Matthew', 'Laura', 'Kevin', 'Jonas')
score <- c(12.5, 9, 16.5, 12, 9, 20, 14.5, 13.5, 8, 19)
attempts <- c(1, 3, 2, 3, 2, 3, 1, 1, 2, 1)
qualify <- c('yes', 'no', 'yes', 'no', 'no', 'yes', 'yes', 'no', 'no', 'yes')

df <- data.frame(name = name, score = score, attempts = attempts, qualify = qualify)
df
```
 
```{r}
str(df)
```
 
```{r}
summary(df)
```
 
```{r}
df[c(3,5),c(1,3)]
```
 
 