---
title: "Session 1 Notebook"
author: "Brian Roepke"
date: "Sept 6, 2020"
output:
  html_document:
    df_print: paged
---
## Assignment

After installing R and RStudio, do the following exercises using RStudio and R  notebook. It's essential to learn the basics of RStudio, how to start a notebook, and save all your work into an HTML document.

**Note**: As of the latest version of R, the assignment operator <- is interchangeable with = operator.  Use either but don't confuse = with == (equality). Single equal is for assignment while double is to test for equality!

In this course we will use R Studio with R Notebooks (**.RMD file**). Notebooks provides interactive code, output and Markdown documentation, which can be exported in multiple file formats. Your code submissions will be HTML. Watch the video: **R Notebook in RStudio** for details.  (Attached is a notebook template along with its html file. Your notebooks should adhere to the sections for your Name, Date, Project Description, formatted using markdown.

### Part 1

**Perform the coding operations (4: R programming basics)  from this link:**

http://www.sthda.com/english/wiki/r-basics-quick-and-easy

**Basic arithmetic operations:**

+ (addition), - (subtraction), * (multiplication), / (division), ^ (exponentiation)

```{r}
7 + 4
7 - 4
7 / 2
7 * 2
```

**Basic arithmetic functions:**

* Logarithms and exponentials: log2(x), log10(x), exp(x)
* Trigonometric functions: cos(x), sin(x), tan(x), acos(x), asin(x), atan(x)
* Other mathematical functions: abs(x): absolute value; sqrt(x): square root.

```{r}
log2(4)
abs(-4)
sqrt(4)
```

**Assigning values to variables:**

```{r}
lemon_price = 2
```

**Basic data types: numeric, character and logical**

```{r}
my_age <- 28
my_name <- "Brian"
# Are you a data scientist?
is_datascientist <- TRUE
```

**Vectors:** a combination of multiple values (numeric, character or logical)

* Create a vector: c() for concatenate
* Case of missing values: NA (not available) and NaN (not a number)
* Get a subset of a vector: my_vector[i] to get the ith element
* Calculations with vectors: max(x), min(x), range(x), length(x), sum(x), mean(x), prod(x): product of the elements in x, sd(x): standard deviation, var(x): variance, sort(x)

```{r}
# Create a numeric vector
friend_ages <- c(27, 25, 29, 26)
mean(friend_ages)
max(friend_ages)
```

**Matrices:** like an Excel sheet containing multiple rows and columns. Combination of multiple vectors with the same types (numeric, character or logical).

* Create and naming matrix: matrix(), cbind(), rbind(), rownames(), colnames()
* Check and convert: is.matrix(), as.matrix()
* Transpose a matrix: t()
* Dimensions of a matrix: ncol(), nrow(), dim()
* Get a subset of a matrix: my_data[row, col]
* Calculations with numeric matrices: rowSums(), colSums(), rowMeans(), colMeans(), apply()

```{r}
rownames = c("row1", "row2", "row3", "row4", "row5")
colnames = c("col1", "col2", "col3")
M = matrix( 
   c(5, 6, 7, 8, 9, 2, 4, 5, 9, 8, 7, 3, 4, 8, 7), 
   nrow=5, 
   ncol=3,
   dimnames = list(rownames, colnames)
   ) 
print (M)
```


**Factors:** grouping variables in your data

* Create a factor: factor(), levels()
* Check and convert: is.factor(x), as.factor(x)
* Calculations with factors:
  + Number of elements in each category: summary(), table()
  + Compute some statistics by groups (for example, mean by groups): tapply()

```{r}
# Create a factor
friend_groups <- factor(c("grp1", "grp2", "grp1", "grp2"))
levels(friend_groups)
```

```{r}
# Compute the mean age by groups
friend_ages <- c(27, 25, 29, 26)
tapply(friend_ages, friend_groups, mean)
```

**Data frames:** like a matrix but can have columns with different types

* Create a data frame: data.frame()
* Check and convert: is.data.frame(), as.data.frame()
* Transpose a data frame: t()
* Subset a data frame: my_data[row, col], subset(), attach() and detach()
* Extend a data frame: $, cbind(), rbind()
* Calculations with numeric data frames: rowSums(), colSums(), rowMeans(), colMeans(), apply()

```{r}
name = c('Nicolas', 'Thierry', 'Bernard', 'Jerome')
age = c(27, 25, 29, 26)
height = c(180, 170, 185, 169)
married = c(TRUE, FALSE, TRUE, TRUE)
df = data.frame(name, age, height, married)
print(df)
```


**Lists:** collection of objects, which can be vectors, matrices, data frames,

* Create a list: list()
* Subset a list
* Extend a list

```{r}
my_family <- list(
  mother = "Veronique", 
  father = "Michel",
  sisters = c("Alicia", "Monica"),
  sister_age = c(12, 22)
  )
# Print
my_family
```

Now on your own. Refer to: https://cran.r-project.org/doc/contrib/Short-refcard.pdf for functions for random, sequence and math operations.

### Part 2

**Create a sequence of numbers from 20 to 50 and find the mean of numbers from 20 to 60 and sum of numbers from 51 to 91**

```{r}
num_seq <- c(20:50)
num_seq
```

**Find the mean of numbers from 20 to 60 and sum of numbers from 51 to 91**

```{r}
mean_seq <- c(20:60)
mean(mean_seq)
```

**Sum of numbers from 51 to 91**

```{r}
sum_seq <- c(51:91)
sum(sum_seq)  
```

### Part 3

**Create a vector which contains 10 random integer values between -50 and +50**

```{r}
# Create the full vector of numbers
full_vector <- c(-51:50)

# Take a sample of 10 of these numbers
sample_vector <- sample(full_vector, 10)
sample_vector
```
