---
title: "Digital Marketing Regression"
output:
  html_document:
    df_print: paged
---

## Linear Regression
Linear regression (or linear model) is used to predict a quantitative outcome variable (y) on the basis of one or multiple predictor variables (x) (James et al. 2014,P. Bruce and Bruce (2017)).

The goal is to build a mathematical formula that defines y as a function of the x variable. Once, we built a statistically significant model, it’s possible to use it for predicting future outcome on the basis of new x values.

The marketing data set included with package datarium is used to create a model for predicting sales units on the basis of the amount of money spent in the 3 advertising medias (youtube, facebook and newspaper)

```{r}
 #install.packages('datarium')
library(dplyr, warn.conflicts = FALSE) # suppress conflicts warning

# Load the data
data("marketing", package = "datarium")

# top 5 observations
head(marketing, 5)
```
## Multiple Linear Regression

```{r}
options(width = 400) # increase width

# correlation
corr <- round(cor(marketing), 2)
corr
```

```{r}
library(corrplot)
```

```{r}
corrplot(corr, method="circle")
```

We see that youtube (.78) and facebook (.58) contribute the most to sales while news paper (0.23) much less.
Create a model with variables youtube and facebook.

```{r}
model <- lm(sales ~ youtube + facebook, data = marketing)

# show model summary
summary(model)
```

easier viewing of coefficients using broom

```{r}
library(broom)
```

```{r}
tidy(model)
```

1. The model equation: sales = 3.50+ 0.045(youtube) + 0.187(facebook). 
2. The intercept (eg: constant) is the expected mean value of Y when all X=0. In this case 3.50
3. All have significant t-statistic values with significant p-values 
4. For a one unit increase in youtube marketing we can expect a 0.045 increase in sales
5. For a one unit increase in facebook marketing we can expect a 0.187 increase in sales

**The Adjusted R-squared (0.896) indicates that about 90% of Sales is explained by this model.**


## Including all variables
Suppose we create a model with all variables

```{r}
model <- lm(sales ~. , data = marketing) # include all variables
summary(model)
```

**From the output, we confirm that the variable newspaper is not statistically significant to the model with a p-value of (0.86)**