---
title: "Facebook EDA"
author: "Brian Roepke"
date: "Oct 4, 2020"
output:
  html_document:
    df_print: paged
---
# Project Description

Perform Exploratory Data Analysis on the provided dataset. Refer to your reading on EDA as well as referenced resources provided in this notebook, e-learning and elsewhere. 
**Your project MUST include meaningful analysis with description, similar to the EDA sample provided.**

```{r}
library(dplyr, warn.conflicts = FALSE) # suppress conflicts warning
library(tidyr)
library(ggplot2)
```

```{r}
fb <- read.csv("Facebook FriendsCSV.csv", skip = 2)

# Clean up some of the columns
fb$Photos <- as.integer(gsub(",","", fb$Photos))
fb$Friends <- as.integer(gsub(",","", fb$Friends))

fb
```

```{r}
str(fb)
```

```{r}
attach(fb)
```

## Data wrangling/munging/transformation
Refers to the process of cleaning, structuring, enriching raw data into a format compatible for down stream processing such as analysis, plotting, etc. Perform the following transformations:

 * Remove Friend field as it is not needed
 * Convert to categorical fields to factors
    * Add AgeGroup (13–17, 18–24, 25–34, 35–44, 45–54, 55-64, 65+)
    * Update Gender values (male=1; female=0, 2=other)
    * Update Sexual Orientation values (1=heterosexual; 0=other)
    * Add InRelationship (1=yes; 0=no) and remove Relationship
    * Update Emp values (employed=1, unemployed=0)
    * Add HasCollege with yes/no: based on Edu where (1=High School or less; 0=college) then remove Edu
    * Update USA values (yes=1; no=0)
    * Add ProfilePic (1=self; 0=other) and remove Profile
    * Add CoverPic (1=self; 0=other) and remove Cover
    * Midwest, Southeast, West are represented. Update their values to be yes/no.
    * But there are observations not represented in these groups. Add a Region column with values: (MidWest, Seast,West, Other) respectively


```{r}
Gender <- as.factor(Gender)
```

