---
title: "Facebook EDA"
author: "Brian Roepke"
date: "Oct 4, 2020"
output: html_document
---
# Project Description

Perform Exploratory Data Analysis on the provided dataset. Refer to your reading on EDA as well as referenced resources provided in this notebook, e-learning and elsewhere. 
**Your project MUST include meaningful analysis with description, similar to the EDA sample provided.**

```{r}
library(dplyr, warn.conflicts = FALSE) # suppress conflicts warning
library(tidyr)
library(ggplot2)
library(car)
library(corrplot)
library(lsr)
library(ggcorrplot)
```

```{r}
fb <- read.csv("Facebook FriendsCSV.csv", skip = 2)
```

```{r}
str(fb)
```

```{r}
names(fb)
```

# Data Prep
Observe the following:

 * X..of..Tags column name was generated due to spaces in the column name. Rename this to Tags
 * The Friend field is not a data field but represents an instance (715 in total), and is not needed
 * Fields with values 0 and 1 are actually categorical (eg: Emp,Profile, Edu, Events, USA, etc..)
 * Gender is categorical ((Male=1; Female=0, 2=Other/Unknown)
 * Fields ‘Photos’ and ‘Friends’ were treated as character when they should be numeric. We can use regular expressions to show which values are causing the data to be treated as chr
 * Remaining fields are continuous

## To Do
1. Rename ‘X of Tags’ to ‘Tags’.
2. Change the datatype from character to integer (Friends, Photos). To do this, need to remove the commas first, using gsub.

```{r}
# Rename ‘X of Tags’ to ‘Tags’.

fb <- rename(fb, "Tags" = "X..of.Tags")
```

```{r}
# Change the datatype from character to integer (Friends, Photos). To do this, need to remove the commas first, using gsub.

fb$Photos <- as.integer(gsub(",","", fb$Photos))
fb$Friends <- as.integer(gsub(",","", fb$Friends))
```

## Verify by looking at the structure again

```{r}
str(fb)
```

## Let’s take a look at the distribution of Age.

```{r}
summary(fb$Age)
```

We see that the min Age is 13, max is 81, average is 24. For plotting, create age groups: 13–17, 18–24, 25–34, 35–44, 45–54, 55-64, 65+

Reference:
https://www.statista.com/statistics/376128/facebook-global-user-age-distribution/

# Data wrangling/munging/transformation
Refers to the process of cleaning, structuring, enriching raw data into a format compatible for down stream processing such as analysis, plotting, etc. Perform the following transformations:

 * Remove Friend field as it is not needed
 * Convert to categorical fields to factors
    * Add AgeGroup (13–17, 18–24, 25–34, 35–44, 45–54, 55-64, 65+)
    * Update Gender values (male=1; female=0, 2=other)
    * Update Sexual Orientation values (1=heterosexual; 0=other)
    * Add InRelationship (1=yes; 0=no) and remove Relationship
    * Update Emp values (employed=1, unemployed=0)
    * Add HasCollege with yes/no: based on Edu where (1=High School or less; 0=college) then remove Edu
    * Update USA values (yes=1; no=0)
    * Add ProfilePic (1=self; 0=other) and remove Profile
    * Add CoverPic (1=self; 0=other) and remove Cover
    * Midwest, Southeast, West are represented. Update their values to be yes/no.
    * But there are observations not represented in these groups. Add a Region column with values: (MidWest, Seast,West, Other) respectively

```{r}
# Remove Friend field as it is not needed

fb$Friend <- NULL
```


```{r}
# Add AgeGroup (13–17, 18–24, 25–34, 35–44, 45–54, 55-64, 65+)

fb <- fb %>% 
  mutate(AgeGroup = case_when(Age >= 13  & Age <= 17 ~ '13–17', 
                              Age >= 18  & Age <= 24 ~ '18-24',
                              Age >= 25  & Age <= 34 ~ '25-34',
                              Age >= 35  & Age <= 44 ~ '35-44',
                              Age >= 45  & Age <= 54 ~ '45-54',
                              Age >= 55  & Age <= 64 ~ '55-64',
                              Age >= 65 ~ '65+'))

fb$AgeGroup <- as.factor(fb$AgeGroup)
```

```{r}
# Update Gender values (male=1; female=0, 2=other)

fb$Gender[fb$Gender == 1] <- 'Male'
fb$Gender[fb$Gender == 0] <- 'Female'
fb$Gender[fb$Gender == 2] <- 'Other'

fb$Gender <- as.factor(fb$Gender)
```

```{r}
# Update Sexual Orientation values (1=heterosexual; 0=other)

fb$Orientation[fb$Orientation == 1] <- 'Heterosexual'
fb$Orientation[fb$Orientation == 0] <- 'Other'

fb$Orientation <- as.factor(fb$Orientation)
```

```{r}
# Add InRelationship (1=yes; 0=no) and remove Relationship

fb <- fb %>% 
  mutate(InRelationship = case_when(Relationship == 1 ~ 'Yes',
                                    Relationship == 0 ~ 'No'))

fb$InRelationship <- as.factor(fb$InRelationship)
fb$Relationship <- NULL
```

```{r}
# Update Emp values (employed=1, unemployed=0)

fb$Emp[fb$Emp == 1] <- 'Employed'
fb$Emp[fb$Emp == 0] <- 'Unemployed'

fb$Emp <- as.factor(fb$Emp)
```

```{r}
# Add HasCollege with yes/no: based on Edu where (1=High School or less; 0=college) then remove Edu

fb <- fb %>% 
  mutate(HasCollege = case_when(Edu == 1 ~ 'High School or Less',
                                Edu == 0 ~ 'College'))

fb$HasCollege <- as.factor(fb$HasCollege)
fb$Edu <- NULL
```

```{r}
# Update USA values (yes=1; no=0)

fb$USA[fb$USA == 1] <- 'Yes'
fb$USA[fb$USA == 0] <- 'No'

fb$USA <- as.factor(fb$USA)
```

```{r}
# Add ProfilePic (1=self; 0=other) and remove Profile

fb <- fb %>% 
  mutate(ProfilePic = case_when(Profile == 1 ~ 'Self',
                                Profile == 0 ~ 'Other'))


fb$ProfilePic <- as.factor(fb$ProfilePic)
fb$Profile <- NULL
```

```{r}
# Add CoverPic (1=self; 0=other) and remove Cover

fb <- fb %>% 
  mutate(CoverPic = case_when(Cover == 1 ~ 'Self',
                              Cover == 0 ~ 'Other'))


fb$CoverPic <- as.factor(fb$CoverPic)
fb$Cover <- NULL
```

```{r}
# Midwest, Southeast, West are represented. Update their values to be yes/no.

fb$MidWest[fb$MidWest == 1] <- 'Yes'
fb$MidWest[fb$MidWest == 0] <- 'No'

fb$Seast[fb$Seast == 1] <- 'Yes'
fb$Seast[fb$Seast == 0] <- 'No'

fb$West[fb$West == 1] <- 'Yes'
fb$West[fb$West == 0] <- 'No'

fb$MidWest <- as.factor(fb$MidWest)
fb$Seast <- as.factor(fb$Seast)
fb$West <- as.factor(fb$West)
```

```{r}
# But there are observations not represented in these groups. Add a Region column with values: (MidWest, Seast,West, Other) respectively

fb <- fb %>% 
  mutate(Region = case_when(MidWest == 'No'  & Seast == 'No' & West == 'No' ~ 'Other',
                           MidWest == 'Yes' ~ 'MidWest',
                           Seast == 'Yes' ~ 'Seast',
                           West == 'Yes' ~ 'West'))

fb$Region <- as.factor(fb$Region)
```

```{r}
str(fb)
```


## Verify Location values
 * create a table showing the new Location values.
 * show top 10 region/location values

```{r}
select(fb, MidWest, Seast, West, Region, USA) %>%
  head(10)
```

```{r}
summary(fb$Region)
```

## Show Top and Bottom observations

```{r}
head(fb)
```

```{r}
tail(fb)
```


# Exploratory Data Analysis (EDA)
 * Show stats, graphs and other exploratory techniques to gain insight.
 * What are your questions? For example, perhaps I’m interested in understanding which variables contribute to the # of Friends. Using this information I can crearte a model that predicts the number of facebook friends.

## Display summary statistics to better understand data distributions by variable.
 * For continuous variables, you have measures of dispersion such as min, max, et…
 * For categorical (factors), you have counts by grouping

```{r}
summary(fb)
```

## Graphs and plots

### Univariant

```{r}
fb %>%
    ggplot(aes(Friends)) +
    geom_histogram(binwidth = 75, color = "blue",fill = "lightblue") +
    geom_vline(xintercept = mean(fb$Friends), lwd = 0.5) +
    geom_vline(xintercept = mean(fb$Friends)-sd(fb$Friends), linetype="dashed") +
    geom_vline(xintercept = mean(fb$Friends)+sd(fb$Friends), linetype="dashed") +
    labs(title = "Distribution of Friends",
         x = "Number of Friends",
         y = "Count") +
    theme_minimal() +
    scale_x_continuous(breaks = seq(0,5000,1000))
```

**Mean number of Friends shown with black line and +/- 1 Standard Deviations**

Here we see that the distribution of number of friends is Right Skewed with a mean of `696`.  We can see around the upper tail of the distribution that a few people have multiple-thousands of friends, but this is not typical. `67%` of the people fall within the dashed lines (`+/- 1 Standard Deviation`) 

```{r}
fb %>%
    ggplot(aes(Posts)) +
    geom_density(color = "red",fill = "salmon") +
    geom_vline(xintercept = median(fb$Posts), lwd = 0.5) +
    labs(title = "Distribution of Posts",
         x = "Number of Posts",
         y = "") +
    theme_minimal() +
    scale_x_continuous(breaks = seq(0,800,100))
```

Here we see the distribution for the number of posts that people made.  The a **median** number of posts is shown in the black line of `38`.  With posts we see an incredible long and flat upper tail showing us a small number of people are posting more than 100 times for this dataset.

```{r}
fb %>%
    ggplot(aes(Age)) +
    geom_histogram(binwidth = 10, color = "black",fill = "gray") +
    geom_vline(xintercept = mean(fb$Age), lwd = 0.5) +
    geom_vline(xintercept = mean(fb$Age)-sd(fb$Age), linetype="dashed") +
    geom_vline(xintercept = mean(fb$Age)+sd(fb$Age), linetype="dashed") +
    labs(title = "Distribution of Age",
         x = "Age",
         y = "Count") +
    theme_minimal() +
    scale_x_continuous(breaks = seq(10,100,10))
```

**Showing Mean Age and +/- 1 Standard Deviations**

For the distribution of people in this dataset according to age, we can see that most people are in their `20's` with the next most common in their `30's`.  


```{r}
fb %>%
    ggplot(aes(Gender,Friends, color = Gender)) +
    geom_boxplot() +
    labs(title = "Distribution of Friends Relative to Gender",
         x = "Gender",
         y = "Friends") +
    theme_minimal() +
    scale_y_continuous(breaks = seq(0,5000,1000)) +
    coord_flip()
```

The number of friends for each Gender is displayed in the box and whisker plots. We can see that the inner quartile ranges for females is larger than for males, showing a little more variation in the number of friends women have vs. men relative to the median number. The median for men and women however, appears to be quite similar to each other, where as the meian for Other is lower.  Note: the number of observations of "Other" in this dataset is relatively small.  

### Multivariate

Plots of multiple variables to look for relationships, etc..

```{r}
ggplot(fb, mapping = aes(x = Posts, y = Friends, color = Gender)) + 
  geom_point(alpha = 0.4) + 
  theme_minimal()
```

For each person in the data set, plotting the number of posts they have vs. the number of friends shows a a high level of concentration below about 1500 friends and below 70-100 posts.  There are a few outliers however that either have a lot of posts and a smaller amout of friends and vice versa.  This dataset does not show entries that have a high amount of both friends and posts. 

```{r}
ggplot(fb, mapping = aes(x = Photos, y = Albums)) + 
  geom_point(color = 'darkgray', alpha = 0.4) + 
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x) + 
  theme_minimal()
```

When looking at the number of photos posted, we can see this coorelates to the number of Albumns a person has as well.  A fairly strong relationship between the two.  This logically makes sense as people are organizging their photos by albumn online.

```{r}
ggplot(fb, mapping = aes(x = Age, y = Posts)) + 
  geom_point(color = 'darkblue', alpha = 0.4) + 
  theme_minimal()
```

Here we see Age on the X axis and Posts on the Y.  We can see that people in their 20's to mid 30's post more than other age groups.There is also a clear visual that the extereme posters fall into this bucket as well. 

```{r}
fb %>% 
    ggplot(aes(AgeGroup)) +
    geom_bar(aes(fill = Gender)) +
    theme_minimal()
```

We can see here that for each Age Group, the number of Males and Females is approximately the same regardless of Age. The exception is "Other", where they appear to be in the 18-24 group, but again.  There is not enough observations in the "Other" gender. 

### Regression Analysis
Now that we’ve conducted EDA, we will create regression model(s) to predict some continuous value. Eg: predict the ‘# of Friends’.

 * Create a multiple linear regression model to predict ‘# of friends’ with only continuous predictor variables (x).
 * Select which x variables should be used in the regression model given correlation analysis. DO NOT split the dataset (use all) and no need to call predict. Only create the model, show its summary and speak to the model equation and what it means.
 * Refer to references and lecture notes

```{r}
fb2 = fb %>% 
  select_if(is.numeric)

 str (fb2)
```

```{r}
corr <- round(cor(fb2), 2)
corr
```

```{r}
corrplot(corr)
```


None of the predictor variables show a strong relationship to the response variable Friends. Of these, Photos (.25), # of Tags (.24), Albums (.23) and Likes (.23) are the strongest. Age (-20) and Children (-16) both have a weak negative correlation. While there appears to be weak correlation over all, these are the strongest and will be further analyzed as to their affect on Friends.

Further there are a number of predictor variables that correlate to each other. For example # of Tags, Photos and Albums are moderately correlated. Similarly Posts and Replies are correlated as is Children and Age. To assess the level of multicollinearity let’s compute the variance inflation factor (VIF)

```{r}
car::vif(lm(Friends ~., data = fb2))
```

We're going to create a simple linear regression between `Age` and `Friends` to get a baseline R-squared value for our dataset.

```{r}
model <- lm(Friends ~ Age, data = fb2)
summary(model)
```

Using Only age we get an R-squared value of 0.039 telling us that only 4% of the variance is explained by age alone.  We will build a Multiple Regression model to see if we can improve on that.

## Multiple regression model

Given the VIF for `Posts`, and `Replies` is greater than `5`, we will eliminate them from the model. Next, after eliminating them, we can see which variables have a *significant* effect on the model. 

```{r}
model <- lm(Friends ~ Age + Photos + Tags + Albums + Children + Likes + Events, data = fb2)
summary(model)
```

After running that multiple regression model, we can see that `Albums`, `Children`, and `Events` are not significant.  We can simplify the model by removing some of these variables. 

```{r}
model <- lm(Friends ~ Age + Photos + Tags + Likes, data = fb2)
summary(model)
```


### Summary
After running the final linear model, we will use a combination of `Age`, `Photots`, `Tags`, and `Likes` to build a predictor for `Friends`.  The R-squared value is 0.136 telling us that 13.6% of the variance in the predicted number of Friends is explained by number number `Photots`, `Tags`, and `Likes` and the person's `Age`.  This is an improvement of over 9% from the simple linear model of just `Friends` and `Age`.  However it's still not a strong predictor.

With this, we can state that this model **is not a sufficient predictor** of the number of Friends someone will have.

**The linear equation is:**

`Y = 827.87677 - 11.93533(Age) + 0.06748(Photos) + 0.10946(Tags) + 0.41635(Likes)`




