---
title: "Facebook EDA"
author: "Brian Roepke"
date: "Oct 4, 2020"
output:
  html_document:
    df_print: tibble
---
# Project Description

Perform Exploratory Data Analysis on the provided dataset. Refer to your reading on EDA as well as referenced resources provided in this notebook, e-learning and elsewhere. 
**Your project MUST include meaningful analysis with description, similar to the EDA sample provided.**

```{r}
library(dplyr, warn.conflicts = FALSE) # suppress conflicts warning
library(tidyr)
library(ggplot2)
```

```{r}
fb <- read.csv("Facebook FriendsCSV.csv", skip = 2)

# Clean up some of the columns
fb$Photos <- as.integer(gsub(",","", fb$Photos))
fb$Friends <- as.integer(gsub(",","", fb$Friends))

fb
```

```{r}
str(fb)
```

## Data wrangling/munging/transformation
Refers to the process of cleaning, structuring, enriching raw data into a format compatible for down stream processing such as analysis, plotting, etc. Perform the following transformations:

 * Remove Friend field as it is not needed
 * Convert to categorical fields to factors
    * Add AgeGroup (13–17, 18–24, 25–34, 35–44, 45–54, 55-64, 65+)
    * Update Gender values (male=1; female=0, 2=other)
    * Update Sexual Orientation values (1=heterosexual; 0=other)
    * Add InRelationship (1=yes; 0=no) and remove Relationship
    * Update Emp values (employed=1, unemployed=0)
    * Add HasCollege with yes/no: based on Edu where (1=High School or less; 0=college) then remove Edu
    * Update USA values (yes=1; no=0)
    * Add ProfilePic (1=self; 0=other) and remove Profile
    * Add CoverPic (1=self; 0=other) and remove Cover
    * Midwest, Southeast, West are represented. Update their values to be yes/no.
    * But there are observations not represented in these groups. Add a Region column with values: (MidWest, Seast,West, Other) respectively

```{r}
# Remove Friend field as it is not needed

fb$Friend <- NULL
```


```{r}
# Add AgeGroup (13–17, 18–24, 25–34, 35–44, 45–54, 55-64, 65+)

fb <- fb %>% 
  mutate(AgeGroup = case_when(Age >= 13  & Age <= 17 ~ '13–17', 
                              Age >= 18  & Age <= 24 ~ '18-24',
                              Age >= 25  & Age <= 34 ~ '25-34',
                              Age >= 35  & Age <= 44 ~ '35-44',
                              Age >= 45  & Age <= 54 ~ '45-54',
                              Age >= 55  & Age <= 64 ~ '55-64',
                              Age >= 65 ~ '65+'))

fb$AgeGroup <- as.factor(fb$AgeGroup)
```

```{r}
# Update Gender values (male=1; female=0, 2=other)

fb$Gender[fb$Gender == 1] <- 'Male'
fb$Gender[fb$Gender == 0] <- 'Female'
fb$Gender[fb$Gender == 2] <- 'Other'

fb$Gender <- as.factor(fb$Gender)
```

```{r}
# Update Sexual Orientation values (1=heterosexual; 0=other)

fb$Orientation[fb$Orientation == 1] <- 'Heterosexual'
fb$Orientation[fb$Orientation == 0] <- 'Other'

fb$Orientation <- as.factor(fb$Orientation)
```

```{r}
# Add InRelationship (1=yes; 0=no) and remove Relationship

fb <- fb %>% 
  mutate(InRelationship = case_when(Relationship == 1 ~ 'Yes',
                                    Relationship == 0 ~ 'No'))

fb$InRelationship <- as.factor(fb$InRelationship)
fb$Relationship <- NULL
```

```{r}
# Update Emp values (employed=1, unemployed=0)

fb$Emp[fb$Emp == 1] <- 'Employed'
fb$Emp[fb$Emp == 0] <- 'Unemployed'

fb$Emp <- as.factor(fb$Emp)
```

```{r}
# Add HasCollege with yes/no: based on Edu where (1=High School or less; 0=college) then remove Edu

fb <- fb %>% 
  mutate(HasCollege = case_when(Edu == 1 ~ 'High School or Less',
                                Edu == 0 ~ 'College'))

fb$HasCollege <- as.factor(fb$HasCollege)
fb$Edu <- NULL
```

```{r}
# Update USA values (yes=1; no=0)

fb$USA[fb$USA == 1] <- 'Yes'
fb$USA[fb$USA == 0] <- 'No'

fb$USA <- as.factor(fb$USA)
```

```{r}
# Add ProfilePic (1=self; 0=other) and remove Profile

fb <- fb %>% 
  mutate(ProfilePic = case_when(Profile == 1 ~ 'Self',
                                Profile == 0 ~ 'Other'))


fb$ProfilePic <- as.factor(fb$ProfilePic)
fb$Profile <- NULL
```

```{r}
# Add CoverPic (1=self; 0=other) and remove Cover

fb <- fb %>% 
  mutate(CoverPic = case_when(Cover == 1 ~ 'Self',
                              Cover == 0 ~ 'Other'))


fb$CoverPic <- as.factor(fb$CoverPic)
fb$Cover <- NULL
```

```{r}
# Midwest, Southeast, West are represented. Update their values to be yes/no.

fb$MidWest[fb$MidWest == 1] <- 'Yes'
fb$MidWest[fb$MidWest == 0] <- 'No'

fb$Seast[fb$Seast == 1] <- 'Yes'
fb$Seast[fb$Seast == 0] <- 'No'

fb$West[fb$West == 1] <- 'Yes'
fb$West[fb$West == 0] <- 'No'

fb$MidWest <- as.factor(fb$MidWest)
fb$Seast <- as.factor(fb$Seast)
fb$West <- as.factor(fb$West)
```

```{r}
# But there are observations not represented in these groups. Add a Region column with values: (MidWest, Seast,West, Other) respectively

fb <- fb %>% 
  mutate(Region = case_when(MidWest == 'No'  & Seast == 'No' & West == 'No' ~ 'Other',
                           MidWest == 'Yes' ~ 'MidWest',
                           Seast == 'Yes' ~ 'Seast',
                           West == 'Yes' ~ 'West'))

fb$Region <- as.factor(fb$Region)
```

```{r}
fb <- rename(fb, "Tags" = "X..of.Tags")
```


```{r}
select(fb, MidWest, Seast, West, Region, USA) %>%
  head(10)
```

```{r}
summary(fb$Region)
```

## Show Top and Bottom observations

```{r}
slice(fb, 10)
```

# Exploratory Data Analysis (EDA)
 * Show stats, graphs and other exploratory techniques to gain insight.
 * What are your questions? For example, perhaps I’m interested in understanding which variables contribute to the # of Friends. Using this information I can crearte a model that predicts the number of facebook friends.

## Display summary statistics to better understand data distributions by variable.
 * For continuous variables, you have measures of dispersion such as min, max, et…
 * For categorical (factors), you have counts by grouping

```{r}
summary(fb)
```

