---
title: "Midterm 1"
author: "Brian Roepke"
date: "October 30, 2020"
output:
  html_document:
    df_print: tibble
---
# Data 104 Midterm

```{r}
# plotting and pipes - tidyverse
library(ggplot2)
library(dplyr, warn.conflicts = FALSE)
options(dplyr.summarise.inform = FALSE) # Suppress summarise info

# text mining library
suppressPackageStartupMessages(library(tidyverse)) # suppress startup message

# date/time libaray
library(lubridate, warn.conflicts = FALSE)
```




```{r}
ga <- read.csv("bq-results.csv")
```


```{r}
ga$date <- as.character(ga$date)
ga$date <- as.Date(ga$date, "%Y%m%d")
ga$browser <- as.factor(ga$browser)
ga$deviceCategory <- as.factor(ga$deviceCategory)
ga$country <- as.factor(ga$country)
ga$region <- as.factor(ga$region)
ga$userID <- as.factor(ga$userID)

# Rescale totalTransactionRevenue by dividing by 1 million
ga$totalTransactionRevenue <- ga$totalTransactionRevenue / 1e+6

# Columns to remove - Not needed for the analysis.  Removing these will make the data a little simpler
ga$hitNumber <- NULL
ga$pagePath <- NULL

str(ga)
```


```{r}
select(ga, date, session, pageviews, transactions,
       totalTransactionRevenue, browser, 
       deviceCategory, country, region) %>%
  summary()
```

```{r}
head(ga)
tail(ga)
```


## Aggregate a View for User Activity Per Session

Values repeat themselves in the table for all variables per session.  So we can group these together to get a more accurate view of each session's activity.

```{r}
ga_aggregate <- ga %>%
  group_by(userID, sessionID) %>%
  distinct()

ga_aggregate
```



```{r}
user_revenue <- ga_aggregate %>%
  filter(!is.na(transactions))

user_revenue
```

```{r}
user_revenue %>%
  ggplot(aes(y = totalTransactionRevenue, x = date)) + 
  geom_col()
```

```{r}
ga_aggregate %>%
  filter(!is.na(transactions)) %>%
  summarise(date, totalTransactionRevenue, pageviews) %>%
  ggplot(aes(x = pageviews, y = totalTransactionRevenue)) + 
  geom_point(alpha = 0.4) + 
  theme_minimal()

```

```{r}
user_revenue %>%
  ggplot(aes(x = date)) + 
  geom_histogram()
```


```{r}
ga_aggregate %>%
    ggplot(aes(browser)) +
    geom_bar() +
    coord_flip() +
    theme_minimal()
```







## Question #1 
What was the average number of product pageviews for users ​who did​ make a purchase?

**Note:** *General calculation: SUM(total_pagesviews_per_user) / COUNT(users)*




```{r}
ga_aggregate %>%
  filter(!is.na(transactions)) %>%
  summarise(total_sessions = sum(session))
```


```{r}
q1 <- ga_aggregate %>%
  filter(!is.na(transactions)) %>%
  group_by(year=floor_date(date, "year"), userID) %>%
  summarise(avg_pageviews = sum(pageviews))

q1
```



## Question #2

What was the average number of product pageviews for users who ​did not​ make a purchase?



## Question #3 
What was the average total transactions per user that made a purchase?

**Note:** *General calculation: SUM (total_transactions_per_user) / COUNT(userID)*


## Question #4
What is the average amount of money spent per session? Here per session is the total of ‘visits’ by user.

**Note:** *General calculation: SUM(total_transactionrevenue_per_user) / SUM(total_visits_per_user)*

## Question #5
What is the total number of transactions generated per browser type ? Results should be in tabular form that shows the aggregated transactions by browser, including those that resulted in 0 transactions.


## Extra Credit

Create a model:
 * Linear Regression - continuous outcome variable
 * Logistic Regression - binary outcome variable;
   * Ex: predicting conversion. Eg: converted = transactions >= 1 is either True (1) or False (0)

