---
title: "2020 Election Analysis using Twitter"
author: "Brian Roepke"
date: "October 25, 2020"
output:
  html_document:
    df_print: tibble
---
# Project Description
Demonstrate how to access Twitter data using rtweet package.
Follow instructions for each section and include text annotation/ interpretation of results!
What can be concluded based on the data investigation?

https://cran.r-project.org/web/packages/rtweet/vignettes/intro.html


```{r}
# twitter library 
library(rtweet)

# plotting and pipes - tidyverse
library(ggplot2)
library(dplyr, warn.conflicts = FALSE)
options(dplyr.summarise.inform = FALSE) # Suppress summarise info

# text mining library
suppressPackageStartupMessages(library(tidyverse)) # suppress startup message

# date/time libaray
library(lubridate, warn.conflicts = FALSE)
```

```{r}
get_token()
```

## Retrieve most recent 3200 timelines of Trump and Biden
 * use: ‘JoeBiden’ and ‘realDonaldTrump’
 * don’t exclude anything
 * save to csv file, so you don’t have to call the api subsequently, in case of long delays
 * read the csv file into a dataframe and use that df for subsequent analysis.
     * You will notice using the View (df) function that the the object structure of the original dataframe after the AP call is different from the one you read in. The original has embedded list objects, whereas the one read in is flattend. Also the date/time fields such as ‘create_at’ is char rather than Twitter’s POSIX time format. Use lubridate datetime functions (eg: as_datetime) to convert.
Show the column names

```{r}
## tmls <- get_timelines(c("JoeBiden", "realDonaldTrump"), n = 3200)
## write_as_csv(tmls, "tweets.csv")
```


```{r}
timelinesdf <- read_twitter_csv("tweets.csv")
```

```{r}
# Convert the date field to a datetime
timelinesdf$created_at <- as_datetime(timelinesdf$created_at)

#Changed some fields to factors for easier manipulation later
timelinesdf$status_id <- as.factor(timelinesdf$status_id)
timelinesdf$screen_name <- as.factor(timelinesdf$screen_name)

# Fix up the reply count field.  It should be a int and NAs set to 0
timelinesdf$reply_count[is.na(timelinesdf$reply_count)] <- 0
timelinesdf$reply_count <- as.integer(timelinesdf$reply_count)
```

```{r}
names(timelinesdf)
```

## Number of Observations for Each Candidate

```{r}
summary(timelinesdf$screen_name)
```

## Show the top and bottom (select by fields shown) and show a summary of ‘created_at’

```{r}
top_and_bottom <- timelinesdf %>%
  select(created_at, screen_name, text)

head(top_and_bottom)
```

```{r}
tail(top_and_bottom)
```


```{r}
summary(top_and_bottom$created_at)
```

Date range for the `3,200` tweets per user received span from `2019-11-09` to `2020-10-18`

## Create a dataframe that filters on just Biden as follows:
 * for timelines past 2020-02-01
 * sorted by favorite_count

Show the top 10 observations (by fields shown) and the count/size

```{r}
df_biden <- timelinesdf %>%
  filter(screen_name == 'JoeBiden' & created_at > '2020-02-01') %>%
  select(created_at, favorite_count, retweet_count, text) %>%
  arrange(desc(favorite_count))

head(df_biden, n = 10L)
```
```{r}
nrow(df_biden)
```

After filtering out the tweets by date, Biden's count dropped to `2,599`, vs. the `3,200` pulled down initially.  

## Retweets, Replies, and Original Tweets Aggregated over Months for Biden


```{r}
df_biden_2 <- timelinesdf %>%
  mutate(mon = floor_date(created_at, 'month')) %>%
  mutate(original = is_retweet==FALSE & is_quote==FALSE & is.na(reply_to_status_id)) %>%
  filter(screen_name == 'JoeBiden' & created_at > '2020-02-01') %>%
  group_by(mon) %>%
  summarise(retweet_count = sum(retweet_count), 
            replies_count = sum(reply_count), 
            orig_count = sum(original))

df_biden_2
```

```{r}
summary(df_biden_2)
```

 * **mon**: dates span from the month of Feb through Oct.  Note - the dates were only through the middle of October and not a full month.  This suggests that activity might increase dramatically for October
 * **retweet_count**: A max of 5.5million retweets.  A staggering number of monthly retweet totals.
 * **replies_count**: This column was empty in the dataset.  After investigating the missing data further, this attribute is only available with the premium API twitter access (see URL below).
 * **orig_count**: This represents the total number of original Tweets by Biden per month. Reasonably consistent with a mean of `227`.
 
 https://stackoverflow.com/questions/47851662/reply-count-attribute-missing-from-tweet-object 

```{r}
df_biden_2 %>%
  ggplot(aes(x = mon)) +
  geom_line(aes(y = retweet_count , color = "retweet"), linetype="dotdash") +
  geom_line(aes(y = replies_count , color = "replies"), linetype="dotdash") +
  geom_line(aes(y = orig_count , color = "original"), linetype="dotdash") +
  scale_color_manual(name = "Legend",
                       values = c("retweet" = "red", 
                                  "replies" = "blue", 
                                  "original" = "black")) +
  scale_y_continuous(trans='log10') +
  theme(
    legend.position = "right",
    plot.title = ggplot2::element_text(face = "bold")) +
  ggplot2::labs(
    x = NULL, y = "Tweets",
    title = "Biden Retweets, Original Tweets, and Replies Over Months",
    subtitle = "01 February 2020 to 09 2020"
  )
```

Plotting the three different values on a log scale, you see precisely those observations.  The number of total retweets is steadily increasing over time, and Biden's original tweet activity is reasonably consistent.

## Comparing the Same Activity, but for Trump

```{r}
df_trump <- timelinesdf %>%
  filter(screen_name == 'realDonaldTrump' & created_at > '2020-02-01') %>%
  select(created_at, favorite_count, retweet_count, text) %>%
  arrange(desc(favorite_count))

df_trump
```
```{r}
nrow(df_trump)
```

Here we can see that during the same period, Trump tweeted `3,178` times vs. Biden's `2,599`  `~800` more tweets in the same period retrieved.

## Show and Plot retweets, replies, and original tweets aggregated over months for Trump

```{r}
df_trump_2 <- timelinesdf %>%
  mutate(mon = floor_date(created_at, 'month')) %>%
  mutate(original = is_retweet==FALSE & is_quote==FALSE & is.na(reply_to_status_id)) %>%
  filter(screen_name == 'realDonaldTrump' & created_at > '2020-02-01') %>%
  group_by(mon) %>%
  summarise(retweet_count = sum(retweet_count), 
            replies_count = sum(reply_count), 
            orig_count = sum(original))

df_trump_2
```

```{r}
summary(df_trump_2)
```

 * **mon**: The dates for Trump are a smaller window regardless of the filter because the API reached the max of `3,200` in `August.`
 * **retweet_count**: A max of `21 million` retweets.  A (`even more`) staggering number of monthly retweet totals.
 * **replies_count**: This column was empty in the dataset.  After investigating the missing data further, this attribute is only available with the premium API Twitter access.
 * **orig_count**: This represents the total number of original Tweets by Trump per month.  While Biden's mean was `227`, Trump's activity was `413` for the window collected.


```{r}
df_trump_2 %>%
  ggplot(aes(x = mon)) +
  geom_line(aes(y = retweet_count , color = "retweet"), linetype="dotdash") +
  geom_line(aes(y = replies_count , color = "replies"), linetype="dotdash") +
  geom_line(aes(y = orig_count , color = "original"), linetype="dotdash") +
  scale_color_manual(name = "Legend",
                       values = c("retweet" = "red", 
                                  "replies" = "blue", 
                                  "original" = "black")) +
  scale_y_continuous(trans='log10') +
  theme(
    legend.position = "right",
    plot.title = ggplot2::element_text(face = "bold")) +
  ggplot2::labs(
    x = NULL, y = "Tweets",
    title = "Trump Retweets, Original Tweets, and Replies Over Months",
    subtitle = "01 February 2020 to 09 2020"
  )
```

We see a much smoother line for the retweets over the shorter period—a more consistent set of values during this time.  Original tweets peeked about September 1st and tapered to October.  But note that this is (1) not a full month, and (2) there are only three data points (months) plotted here.

## Plot the frequency of tweets for each user

### by days

```{r}
## plot the frequency of tweets for each user over time
timelinesdf %>%
  dplyr::group_by(screen_name) %>%
  ts_plot("days", trim = 0L) +
  ggplot2::geom_line() +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    legend.title = ggplot2::element_blank(),
    legend.position = "right",
    plot.title = ggplot2::element_text(face = "bold")) +
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Frequency of Tweets by Candidate by Day",
    subtitle = "Last 3,200 Tweets per candidate retreived",
    caption = "\nSource: Data collected from Twitter's REST API via rtweet"
  )
```

This plot visualizes the point that was made above with the compressed window of time.  Trump's tweet volume hit the `3,200` tweet limit in August, whereas Biden overall has had much less activity, and therefore we see a more extended timeline.

This visual also nicely shows the volume of activity per day of Biden vs. Trump.  While it has increased for Biden during the period we have for Trump, we can see the overall per day tweets from Trump is about `5-8x` that of Biden.

### by weeks

```{r}
## plot the frequency of tweets for each user over time
timelinesdf %>%
  dplyr::group_by(screen_name) %>%
  ts_plot("weeks", trim = 0L) +
  ggplot2::geom_line() +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    legend.title = ggplot2::element_blank(),
    legend.position = "right",
    plot.title = ggplot2::element_text(face = "bold")) +
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Frequency of Tweets by  Candidate by Week",
    subtitle = "Last 3,200 Tweets per candidate retreived",
    caption = "\nSource: Data collected from Twitter's REST API via rtweet"
  )
```

When viewing the same data aggregated by Week, we can see a little bit of the noise taken out.  Trump was consistently tweeting weekly more than Biden. 

### By Month (Adding Tweets vs. Retweets)

```{r}
## plot the frequency of tweets for each user over time
timelinesdf %>%
  dplyr::group_by(screen_name, is_retweet) %>%
  ts_plot("month", trim = 0L) +
  ggplot2::geom_line() +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    legend.position = "right",
    plot.title = ggplot2::element_text(face = "bold")) +
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Frequency of Tweets by Candidate",
    subtitle = "Last 3,200 Tweets per candidate retreived",
    caption = "\nSource: Data collected from Twitter's REST API via rtweet"
  )
```

This graph shows the data aggregated by Month, with the addition/differentiation of showing the tweets (*solid line*) vs. retweets (*dashed line*).  This shows how Trump is a much more active retweeter vs. Biden, by a significant margin.

## Get number of followers for Biden and Trump

```{r}
lu_users <- lookup_users(c("JoeBiden", "realDonaldTrump"))

lu_users %>%
  select(screen_name, followers_count)

```
Trump's followers are `87 million` vs. Biden's `11 million`. Followers in social media signify how much reach someone has; however, Twitter is an open platform, so reach/read doesn't equate to how many people are exposed to the content.

## Search 5000 tweets with
 * the hashtag #2020Election
 * exclude retweets
 * save tweets to csv file

```{r}
# election_hastag <- search_tweets("#2020Election", n = 5000, include_rts = FALSE)
# write_as_csv(election_hastag, "hashtag.csv")
```

```{r}
election_hastag <- read_twitter_csv("hashtag.csv")
```

```{r}
# Convert the date field to a datetime
election_hastag$created_at <- as_datetime(election_hastag$created_at)

#Changed some fields to factors for easier manipulation later
election_hastag$status_id <- as.factor(election_hastag$status_id)
election_hastag$screen_name <- as.factor(election_hastag$screen_name)
election_hastag$location <- as.factor(election_hastag$location)
election_hastag$urls_expanded_url <- as.factor(election_hastag$urls_expanded_url)

# Fix up the reply count field.  It should be a int and NAs set to 0
election_hastag$reply_count[is.na(election_hastag$reply_count)] <- 0
election_hastag$reply_count <- as.integer(election_hastag$reply_count)
```

## Show summary of creation date

```{r}
summary(election_hastag$created_at)
```

## Plot hourly timeline of tweets

```{r}
election_hastag %>%
  ts_plot("hours", trim = 0L) +
  ggplot2::geom_line() +
  ggplot2::theme(
    plot.title = ggplot2::element_text(face = "bold")) +
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Frequency of Tweets with #2020 Hashtag",
    subtitle = "Oct 16, 2020 to Oct 19, 2020",
    caption = "\nSource: Data collected from Twitter's REST API via rtweet"
  )
```

This time series is plotting `5,000` tweets on an hourly basis.  The X-axis scale shows the daily spacing, and the peaks and valleys of the data show a frequency of activity increasing around 8:00 am Eastern Time and then maintaining a push through 8:00 pm eastern.  The final day of activity does have a slightly higher peak. 

This consistency probably suggests that most of these tweets are from the same region of the world.  If they were global, we wouldn't see the consistent up and down pattern we see here. Most likely, the traffic is US based, and we can explore this next.

**Note**: *Twitter returns time in UTC format.  Therefore UTC 12:00 pm will be 8:00 am eastern time.*
https://developer.twitter.com/en/docs/twitter-ads-api/timezones 

```{r}
# Clean up some of the obvious data discrepancies.  This is an impossible task to fully clean up, but for the purpose of the "top 10" results we can make it more accurate.
election_hastag$location[election_hastag$location == "Washington, DC"] <- "Washington D.C."
election_hastag$location[election_hastag$location == "Washington, D.C."] <- "Washington D.C."
election_hastag$location[election_hastag$location == "United States of America"] <- "USA"
election_hastag$location[election_hastag$location == "California"] <- "California, USA"
election_hastag$location[election_hastag$location == "Texas"] <- "Texas, USA"
```


```{r}
loc <- election_hastag %>%
  select(location) %>%
  group_by(location) %>%
  drop_na() %>%
  tally(sort = TRUE, name = "count") %>%
  head(n = 15L)
```

```{r}
loc[1:10,] %>% 
  ggplot() +
  geom_col(aes(x = reorder(location, count), y = count)) + 
  coord_flip() +
  ggplot2::labs(
    x = "Count", 
    y = "Location",
    title = "Where Twitter users are from - Unique Locations",
    caption = "\nSource: Data collected from Twitter's REST API via rtweet"
  )
```

As suspected, The top 10 tweets show US-based locations. Position `15` shows `Melbourne` as the location, either Australia or Florida, which has the same spelling as the Australian city.  Of course, this is to be expected because it's a US election; however, the US affects the global economy, so it tends to be watched by many people around the world.

## Show top 10 most frequently shared link (urls_expanded_url)


```{r}
election_hastag %>%
  count(urls_expanded_url, sort = TRUE) %>%
  drop_na() %>%
  head(n = 10L)
```


## Show top 3 most retweeted tweet (retweet_count)

```{r}
election_hastag %>%
  select(created_at, retweet_count, screen_name, text) %>%
  arrange(desc(retweet_count)) %>%
  head(n = 3L)
```

## Show top 3 most favorited tweet (hint: favorite_count)

```{r}
election_hastag %>%
  select(created_at, favorite_count, screen_name, text) %>%
  arrange(desc(favorite_count)) %>%
  head(n = 3L)
```

## Show top 10 tweeters

```{r}
election_hastag %>%
  count(screen_name, sort = TRUE) %>%
  head(n = 10L)
```

## What is the top tweeter saying?

```{r}
# top_tweeter <- get_timelines("usaguyme", n = 100)
# write_as_csv(top_tweeter, "top_tweeter.csv")
```

```{r}
top_tweeter <- read_twitter_csv("top_tweeter.csv")

# Convert the date field to a datetime
top_tweeter$created_at <- as_datetime(top_tweeter$created_at)

#Changed some fields to factors for easier manipulation later
top_tweeter$screen_name <- as.factor(top_tweeter$screen_name)
```

```{r}
top_tweeter %>%
  select(created_at, retweet_count, text) %>%
  arrange(desc(retweet_count)) %>%
  head(n = 20L)
```

**Note 1**: I selected the second users from the list.  The first one was essentially a bot that tweeted out the number of followers each of the candidates had.  Not interesting.  

What we can see from the list is that two of this user's recent tweets went "viral: one generating `~45k` retweets and another `~14.5k` retweets.  The rest are relatively benign.  The text for the two most retweeted tweets are:

 * *"The salary of a U.S. Senator is $174,000 per year. This is Joe Biden’s house.... seems legit https://t.co/DtD0DzXlrY"*

 * *"After corruptly blocking readers from accessing the New York Post’s expose on Hunter and Joe Biden, Twitter is now throttling circulation of a Fox News story detailing Biden’s business ties in communist China. https://t.co/CIaE1raYHP"*

```{r}
top_profile <- lookup_users("usaguyme")
create_date <- top_profile$account_created_at
create_date
```

**Note 2**: This user profile, which made it to the top of this hashtag list, was created on October 16th, 2008.  I suspected that this might be a bot account, but judging by the fact that the account is 12 years old, it's probably a human. While this profile only has 129 followers, we can see here, just by picking up the right hashtag, a user can get themselves into the conversation.


