---
title: "Assignment 13"
author: "Brian Roepkee"
date: "Dec 6, 2020"
output:
  html_document:
    df_print: paged
---
# Sentiment Analysis

1) Read Ch 2 on sentiment analysis as well as sentiment analysis section of Ch 9 to understand how to approach lexicon-based sentiment analysis using tidytext

2) Using the twitter data from your previous assignment with Biden and Trump tweets , perform sentiment analysis using the afinn lexicon.  The tweets dataset has been attached..

3) Compare sentiment between each candidate with full analysis and visualizations.  Show most positive /negative words between each candidate. You can use bar blots and/or comparison.cloud() to group words by positive/negative sentiment.  See ch 2 of your reading and sample notebook The sample notebook has specific instructions for transformations.

4) Extra Credit: (+ 5) points.  If you do this, add an extra credit section as a header. Perform sentiment analysis using bing lexicon.

## Load Libraries and Data

```{r message=FALSE, warning=FALSE}
# twitter library 
library(rtweet)

# plotting and pipes
library(tidyverse)
library(ggplot2)
library(dplyr)
library(stringr)
library(tidyr)

# text mining library
library(tm)
library(tidytext)
library(wordcloud)
library(reshape2)
library(textstem)

# date/time library
library(lubridate)
```


```{r}
df <- read.csv("candidates.csv")
```


```{r}
# Convert the date field to a datetime
df$created_at <- as_datetime(df$created_at)

#Changed some fields to factors for easier manipulation later
df$user_id <- as.factor(df$user_id)
df$status_id <- as.factor(df$status_id)
df$screen_name <- as.factor(df$screen_name)

# Fix up the reply count field.  It should be a int and NAs set to 0
df$reply_count[is.na(df$reply_count)] <- 0
df$reply_count <- as.integer(df$reply_count)
```


```{r}
head(df)
```


```{r}
tail(df)
```

### Filter out Retweets

Validate the number of rows before and after removing retweets.

```{r}
nrow(df)
```


```{r}
df <- df %>%
  filter(is_retweet == FALSE)
head(df)
```

Show the total number of rows in this dataset.

```{r}
nrow(df)
```

### Add Doc_Id incrementing per Row

```{r}
df = df %>%
  mutate(doc_id = paste0("doc", row_number())) %>%
  select(doc_id, everything())
head(df)
```

### Add a Column for Text Length

```{r}
df$text_len <- str_count(df$text)
```

# EDA

Simplify the Data Frame to just the target fields we'll be using throughout the analysis.

```{r}
df_select <- df %>%
  select(doc_id, user_id, status_id, screen_name, created_at, text, text_len)
```

## Tweet Length
Number of tweets by each candidate

```{r}
summary(df_select$text_len)
```

Tweets for the dataset range from `7` characters to `316` with a mean of `172`


```{r}
df_select %>%
  ggplot(aes(x = text_len, fill = screen_name)) +
  geom_histogram(alpha = .5, color = "darkgray", bins = 33) +
  theme_minimal()
```

Trumps tweets are a little more right skewed than Biden's, which appear to be a little left skewed (tend to be longer).  Both candidates have a very high number of short tweets (~20 characters)

### Total Number of Tweets

```{r}
df_select %>%
  group_by(screen_name) %>%
  summarise(total = n(), .groups = "keep") %>%
  ggplot(aes(x = screen_name, y = total, fill = screen_name)) +
  geom_col() + 
  theme_minimal() +
  geom_text(aes(label = total), position = position_stack(vjust = 0.5))
```

After removing Retweets, Trumps total dropped to only `1,733`, while Biden's fell much less to `3,032`.  Trump Retweets more often than Biden.

### Tweets Over Time

```{r}
df_select %>%
  dplyr::group_by(screen_name) %>%
  ts_plot("days", trim = 0L) +
  ggplot2::geom_line() +
  ggplot2::theme_minimal() +
  ggplot2::theme(
    legend.title = ggplot2::element_blank(),
    legend.position = "right",
    plot.title = ggplot2::element_text(face = "bold")) +
  ggplot2::labs(
    x = NULL, y = NULL,
    title = "Frequency of Tweets by Candidate by Day"
  )
```

Based on the way this dataset was collected (last 3200 tweets per candidate), the data from Trump is compressed to dates starting around August.  This is due to the fact that he tweets more regularly than Biden, hitting the 3,200 tweet limit sooner.

# Clean and Prepare the Text for Analysis

Get rid of various useless text like URLS and shortened URLs.  These appear frequently in the text and skew results.

### Clean Twitter specific items
https://stackoverflow.com/questions/31348453/how-do-i-clean-twitter-data-in-r

```{r}
df_select$text = gsub("[ \t]{2,}", " ", df_select$text)
df_select$text = gsub("(RT|via)((?:\\b\\W*@\\w+)+)", "", df_select$text)
df_select$text = gsub("@\\w+", " ", df_select$text)
df_select$text = gsub("[[:digit:]]", " ", df_select$text)
df_select$text = gsub("http\\w+", " ", df_select$text)
df_select$text = gsub("^\\s+|\\s+$", " ", df_select$text)
```

### Remove Whitespace and Lemmatize

```{r}
df_select$text = stripWhitespace(df_select$text)
df_select$text = lemmatize_strings(df_select$text)
```


### Unnest_Tokens()

Create a new column with each word on it's own row.

```{r}
tidy_df <- df_select %>%
  unnest_tokens(word, text)
```

### Stop Word Removal
remove stop words and custom stop words from the results.

Note: Removing Trump because it's a proper name but also an verb which was marked as positive and appeared in the top results for both candidates. 

```{r}
custom_stop_words <- bind_rows(tibble(word = c("trump"), lexicon = c("custom")), stop_words)

tidy_df <- tidy_df %>%
  anti_join(custom_stop_words, by = "word")
```

### Validate the New number of Rows
Dramatically larger now that each word from text is in it's own row.

```{r}
nrow(tidy_df)
```

After unnesting the words, each word of the tweet is on a separate line. The following is an example. 

```{r}
tidy_df %>%
  filter(status_id == "x1319032499456987136") %>%
  head(n=10)
```


# Sentiment Analysis

## Bing Sentiment Lexicon
Using the Bing Lexicon from Bing Liu and collaborators, adds the column "Sentiment" and mark each word as positive or negative. 

https://www.cs.uic.edu/~liub/FBS/sentiment-analysis.html

```{r}
bing_df <- tidy_df %>%
  inner_join(get_sentiments("bing"), by = "word")
```


```{r}
bing_df %>%
  group_by(screen_name, sentiment) %>%
  summarise(count = n(), .groups = "keep")
```



### AFINN scoring Lexicon
AFINN from Finn Årup Nielsen, adds the `value` column, with a numeric representation of how positive, or negative the word is.  The AFINN lexicon measures sentiment with a numeric score between -5 and 5

http://www2.imm.dtu.dk/pubdb/pubs/6010-full.html 

```{r}
afinn_df <- tidy_df %>%
  inner_join(get_sentiments("afinn"), by = "word")

head(afinn_df)
```

```{r}
afinn_df %>%
  ggplot(aes(x = value, fill = "#F8766D")) +
  geom_histogram(bins = 10, show.legend = FALSE) +
  scale_x_continuous(breaks = c(-5, -3, -1, 1, 3, 5)) +
  theme_minimal()
```

For the dataset overall, there is a slight left-skew showing there is a greater concentration of words with positive values.  There are very few in the high and low values (-4,-5, 4, 5).

### NRC Sentiment Lexicon
NRC from Saif Mohammad and Peter Turney. The NRC Emotion Lexicon is a list of English words and their associations with eight basic emotions as well as positive and negative sentiment.

One thing to note, single words can have multiple emotions

```{r}
nrc_df <- tidy_df %>%
  inner_join(get_sentiments("nrc"), by = "word")
```

Total counts for all 8 emotions and 2 sentiments.

```{r}
nrc_df %>%
  group_by(sentiment) %>%
  summarise(total = n(), .groups = "keep") %>%
  arrange(desc(total))
```

## Inspect Top Words Per Candidate
Using various methods, inspect what words are most frequently used, per candidate, proportions of negative and positive words, and trend over time. 

### Top Word Counts (BING)

```{r message=FALSE, warning=FALSE}
bing_df %>%
  count(word, sort = TRUE, sentiment) %>%

  group_by(sentiment) %>%
  top_n(15) %>%
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  theme_minimal() +
  labs(x = "Contribution to sentiment",
       y = NULL)
```

When looking at data as a whole, we can see the top negative word is `crisis` and positive is `win`.  Next, we'll split these out by candidate. 

### Top Word Count Per Candidate (BING)

```{r message=FALSE, warning=FALSE}
bing_df %>%
  count(word, sort = TRUE, screen_name, sentiment) %>%

  group_by(screen_name, sentiment) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(vars(screen_name, sentiment), scales = "free_y") +
  theme_minimal() +
  labs(x = "Contribution to sentiment",y = NULL)
```

**Negative**: Biden has the top word `crisis` in his list, where Trump doesn't.  Biden uses `threat`, `hate`, `fear`, and `lie` in his negative list where Trump has `fake`, `crime`, `radical`, `corrupt`, and `crazy`.

**Positive**: Both candidates show `win` as the top word and `support`, `strong`, `honor`, and `protect` in their top 10. Biden uses `promise`, `safe`, `protect`, and `love`.  Trump uses `endorsement`, `fast`, `congratulations`, `incredible` and `happy`.

### Overall Top Words (BING)

```{r message=FALSE, warning=FALSE}
bing_df %>%
  count(word, sort = TRUE, sentiment, screen_name) %>%

  group_by(screen_name) %>%
  top_n(30) %>%
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = TRUE) +
  facet_wrap(~screen_name, scales = "free_y") +
  theme_minimal() +
  labs(x = "Contribution to sentiment", y = NULL)
```

Looking at the mix of top 30 words and whether they're positive or negative.  Both candidates are about the same.  There is a mix of positive and negative words throughout.   Biden tends to have a little more postive words clustered at the top where Trump is more mixed.

### Top Words Sorted by AFINN Score

```{r message=FALSE, warning=FALSE}
afinn_df$color <- ifelse(afinn_df$value < 0, "Negative","Positive")

afinn_df %>%
  count(word, sort = TRUE, screen_name, value, color) %>%
  group_by(screen_name) %>%
  top_n(30) %>%
  ungroup() %>%
  mutate(word = reorder(word, value)) %>%
  
  ggplot(aes(value, word, fill = color)) +
  geom_col(show.legend = TRUE) +
  facet_wrap(~screen_name, scales = "free_y") +
  theme_minimal() +
  labs(x = "AFINN Sentiment Score", y = NULL)
```
Top 30 words, sorted by their AFINN score, a scale of -5 to 5. Biden used slightly more positive words in his top 30. Both seem generally balanced when sorted this way.

### Overall Sentiment per Candidate (AFINN)

```{r}
afinn_df %>%
  count(word, sort = TRUE, screen_name, value) %>%
  group_by(screen_name) %>%
  summarize(avg = mean(value * n), .groups = "keep") %>%
  
  mutate(screen_name = reorder(screen_name, avg)) %>%
  ggplot(aes(avg, screen_name, fill = screen_name, label = avg)) +
  geom_col() +
  labs(x = "Mean sentiment Score", y = NULL) +
  theme_minimal() +
  geom_text(position = position_dodge(width = 0.9), vjust = -0.5) +
  coord_flip()
```


The MEAN sentiment score is a mean calculation using the AFINN lexicon.  As a remember this scale can go from -5 to +5.  The calculation takes into account the AFINN sentiment score multiplied by the number of times that work occurs and then calculates the MEAN.  We see that Biden's is approximately +1 while Trumps is just over 0. 

### Sentiment of Tweets over Time (AFINN)

```{r}
plot_df2 <- afinn_df %>%
  filter(created_at > "2020-08-01") %>%
  mutate(mon = floor_date(created_at, "day")) %>%
  group_by(screen_name, mon) %>%
  summarize(value = mean(value), .groups = 'keep')

plot_df2$color <- ifelse(plot_df2$value < 0, "negative","positive")

ggplot(plot_df2, aes(mon, value, fill = color)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~screen_name, ncol = 1, scales = "free_x") +
  labs(x = NULL, y = "Mean AFINN Sentiment Score") +
  theme_minimal()
```

Since August, showing the frequency of sentiment for each candidate.  Using the mean sentiment score we can see how each candidate tweeted on a daily basis.  Trumps earlier timeline tended to be more negative, while both candidates were more positive starting mid September. 

### Top Emotions Per Candidate

```{r}
nrc_biden <- nrc_df %>%
  filter(screen_name == "JoeBiden") %>%
  mutate(sum = n()) %>%
  group_by(sentiment, word, sum) %>%
  summarise(avg = mean(n() / sum), .groups = "keep") %>%
  group_by(sentiment) %>%
  summarise(mean = sum(avg), .groups = "keep") %>%
  arrange(desc(mean))

nrc_trump <- nrc_df %>%
  filter(screen_name == "realDonaldTrump") %>%
  mutate(sum = n()) %>%
  group_by(sentiment, word, sum) %>%
  summarise(avg = mean(n() / sum), .groups = "keep") %>%
  group_by(sentiment) %>%
  summarise(mean = sum(avg), .groups = "keep") %>%
  arrange(desc(mean))
```

```{r}
nrc_biden$screen_name <- "JoeBiden"
nrc_trump$screen_name <- "realDonaldTrump"
nrc_final <- rbind(nrc_trump, nrc_biden)
```

```{r}
ggplot(nrc_final, aes(x = mean, y = sentiment, fill = screen_name)) + 
  geom_col(position = "dodge") + 
  theme_minimal() +
  labs(
    x = "Count", y = "NRC Sentiment/Emotion",
    title = "Mean NRC Sentiment",
    subtitle = "Mean Count of Words in Each Category"
  ) +
  theme(
    legend.title = element_blank(),
    legend.position = "right",
    plot.title = element_text(face = "bold"))
```



### Find most positive and negative individual messages by user (AFINN)

#### Most Positive Biden Messages

```{r}
afinn_df %>%
  filter(screen_name == "JoeBiden") %>%
  group_by(doc_id) %>%
  summarize(total_value = sum(value), word_count = n(), .groups = "keep") %>%
  arrange(desc(total_value))
```

```{r}
df %>%
  filter(doc_id == "doc2737" | doc_id == "doc2894") %>%
  select(text)
```

#### Most Negative Biden Messages

```{r}
afinn_df %>%
  filter(screen_name == "JoeBiden") %>%
  group_by(doc_id) %>%
  summarize(total_value = sum(value), word_count = n(), .groups = "keep") %>%
  arrange(total_value)
```


```{r}
df %>%
  filter(doc_id == "doc2504" | doc_id == "doc623") %>%
  select(text)
```


#### Most Positive Trump Messages

```{r}
afinn_df %>%
  filter(screen_name == "realDonaldTrump") %>%
  group_by(doc_id) %>%
  summarize(total_value = sum(value), word_count = n(), .groups = "keep") %>%
  arrange(desc(total_value))
```

```{r}
df %>%
  filter(doc_id == "doc3173" | doc_id == "doc3708") %>%
  select(text)
```

#### Most Negative Trump Messages

```{r}
afinn_df %>%
  filter(screen_name == "realDonaldTrump") %>%
  group_by(doc_id) %>%
  summarize(total_value = sum(value), word_count = n(), .groups = "keep") %>%
  arrange(total_value)
```

```{r}
df %>%
  filter(doc_id == "doc3190" | doc_id == "doc4219") %>%
  select(text)
```

# Word Clouds

## Overall Word Cloud

```{r message=FALSE, warning=FALSE}
bing_df %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 150))
```

## Word Clouds for Each Candidate

### Biden

```{r message=FALSE, warning=FALSE}
bing_df %>%
  filter(screen_name == "JoeBiden") %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 150))
```

### Trump

```{r message=FALSE, warning=FALSE}
bing_df %>%
  filter(screen_name == "realDonaldTrump") %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 150))
```


```{r message=FALSE, warning=FALSE}
bing_df %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("#F8766D", "#00BFC4"), max.words = 200)
```

# Checking for Top-Negation Word Pairs
The method used here doesn't take into account negation of words like no, and never preceding words.  We can quckly check the top word pairs to see how they occur in our text.


```{r}
bigrams <- df_select %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2)
```


```{r}
bigram_counts <- bigrams %>%
  count(bigram, sort = TRUE) %>%
  separate(bigram, c("word1", "word2"), sep = " ")

bigram_counts
```


```{r}
negate_words <- c("not", "without", "no", "can't", "don't", "won't")


bigram_counts %>%
  filter(word1 %in% negate_words) %>%
  count(word1, word2, wt = n, sort = TRUE) %>%
  inner_join(get_sentiments("afinn"), by = c(word2 = "word")) %>%
  mutate(contribution = value * n) %>%
  group_by(word1) %>%
  slice_max(abs(contribution), n = 10) %>%
  ungroup() %>%
  mutate(word2 = reorder_within(word2, contribution, word1)) %>%

  ggplot(aes(contribution, word2, fill = contribution > 0)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~word1, scales = "free", nrow = 2) +
  scale_y_reordered() +
  labs(x = "Sentiment value * # of occurrences",
       y = "Words preceded by a negation")
```

Both `not` and `no` have a high proportion of positive sentiment scored wored associated with them.  Further techniques could take these into account such as removing them via the custom stop words or potentially other methods. However, for this exercise, they will remain in the model. 


