---
title: "Assignment 6A"
author: "Brian Roepke"
date: "October 18, 2020"
output:
  html_document:
    df_print: paged
---
## Assignment 6A: Google Studio HOA & R Programming

```{r}
library(dplyr, warn.conflicts = FALSE)
library(readxl)
library(tidyr)
```

```{r}
df <- read_excel('SocialMediaRevenue.xlsx', 'Dataset1') 
```

```{r}
df <- df %>%
  rename(social_media = 'Social Network') %>% 
  rename_all(tolower) %>%
  drop_na()
```

```{r}
df
```

### 2) Perform data translations:
- add these columns. (hint: use mutate) Observe values are percentages!
    - `session_share` = `sessions` / `sum(sessions)` 
    - `sales_share` = `transactions` /  `sum(transactions)` 
    - `revenue_share` = `revenue` / `sum(revenue)`
- Rename transactions to `sales`
- Then sort by `session_share` in descending order.

```{r}

```


### 3) Show the structure, contents and summary of the transformed data frame.

```{r}

```


### 4) Plot `session_share` vs `sales_share` by `social_media.`
This plot shows the relationship between `sessions` vs `sales` (as a percentage of total) for each social media, even those which did not produce any sales. That is, do not perform any filtering.

```{r}

```

Describe insights gained from the visualization.


### 5) Create the same plot above, but this time, filter the data set to only show observations with `sales > 0`. Following is an example

```{r}

```

